---
title: "R1 project"
subtitle: "Group 10"
author: 
- Christopher Comora
- Nataliya Peshekhodko 
- Zhuoning Liu
date: '`r format(Sys.Date(), "%Y-%B-%d")`'

output:
  pdf_document: default
  html_document: default
---

```{r, echo = FALSE}
set.seed(10)
```

## 1. General Procedure
```{r}
n=10
lambda = 1
```
1.a Generate a random sample of size `r n` with parameter $\lambda$=`r lambda`. 
```{r}
data = rpois (n, lambda)
```
1. b Create and save a 95% CI for the mean of the population using the approximate
confidence interval 1.
```{r}
lambda_hat = mean (data)
approx_var = var (data)

lower_bound_ci1 = lambda_hat + qnorm(0.025)*sqrt(approx_var/length(data))
upper_bound_ci1 = lambda_hat - qnorm(0.025)*sqrt(approx_var/length(data))
ci_1 <- c(lower_bound_ci1, upper_bound_ci1)
ci_1
```
1. c Create and save a 95% CI for the mean of the population using the approximate
confidence interval 1.
```{r}
lower_bound_ci2 = lambda_hat + qnorm(0.025)*sqrt(lambda_hat/length(data))
upper_bound_ci2 = lambda_hat - qnorm(0.025)*sqrt(lambda_hat/length(data))
ci_2 <- c(lower_bound_ci2, upper_bound_ci2)
ci_2
```
## 2. Repeat the general procedure `N=5000` times

## 3. Summary about different method execution

## 4. Repeat steps above for diffrent sample sizes 

## 5. Advantages and disadvantages of different types of interval 

## 6. Apply the 4 different methods to the horse-kick death data and report the 95% confidence intervals obtained by each method.

