---
title: "R1 project"
subtitle: "Group 10"
author: 
- Christopher Comora
- Nataliya Peshekhodko 
- Zhuoning Liu
date: '`r format(Sys.Date(), "%Y-%B-%d")`'

output:
  pdf_document: default
  html_document: default
---

```{r, echo = FALSE}
library(knitr)
set.seed(10)
```

## 1. General Procedure
```{r}
n=10
lambda = 1
B = 1000
alpha = 0.05
```
1.a Generate a random sample of size `r n` with parameter $\lambda$=`r lambda`. 
```{r}
data = rpois (n, lambda)
```
1. b Create and save a 95% CI for the mean of the population using the approximate
confidence interval 1.
```{r}
lambda_hat = mean (data)
approx_var = var (data)

lower_bound_ci1 = lambda_hat - qnorm(1-alpha/2)*sqrt(approx_var/length(data))
upper_bound_ci1 = lambda_hat + qnorm(1-alpha/2)*sqrt(approx_var/length(data))
ci_1 <- c(lower_bound_ci1, upper_bound_ci1)
ci_1
```

1. c Create and save a 95% CI for the mean of the population using the approximate
confidence interval 2.
```{r}
lower_bound_ci2 = lambda_hat - qnorm(1-alpha/2)*sqrt(lambda_hat/length(data))
upper_bound_ci2 = lambda_hat + qnorm(1-alpha/2)*sqrt(lambda_hat/length(data))
ci_2 <- c(lower_bound_ci2, upper_bound_ci2)
ci_2
```

1.d Create and save 95% CIs for the mean of the population using the parametric
bootstrap method. ($B$ = `r B`)
```{r}
bootData <- replicate (n=B, expr = rpois (n, lambda_hat))
means<-colMeans(bootData)
c(quantile (means, 0.025), quantile (means, 0.975))
```

1.e Create and save 95% CIs for the mean of the population using the nonparametric
bootstrap method. ($B$ = `r B`)
```{r}
lambda_vec <- rep(NA, B)
for(i in 1:B){
sel <- sample(1:n, n, replace=TRUE)
  bootstrap_x <- data[sel]
  lambda_vec[i] <- mean(bootstrap_x) }
CI_b <- quantile(lambda_vec,c(0.025,0.975))
CI_b
```

## 2. Repeat the general procedure `N=5000` times
```{r}
N = 5000

means <- c()
lower_bound <- c()
upper_bound <- c()

for (j in 1:N){
  data = rpois (n, lambda)
  means[j] = mean (data) 
  lower_bound[j] = means[j] - qnorm(1-alpha/2)*sqrt(var(data)/length(data))
  upper_bound[j] = means[j] + qnorm(1-alpha/2)*sqrt(var(data)/length(data))  
}
ci_1 <- data.frame ("lower_bound" = lower_bound, "upper_bound" = upper_bound)


lower_bound <- c()
upper_bound <- c()
for (j in 1:N) {
  lower_bound[j] = means[j] - qnorm(1-alpha/2)*sqrt(means[j]/length(data))
  upper_bound[j] = means[j] + qnorm(1-alpha/2)*sqrt(means[j]/length(data))
}
ci_2 <- data.frame ("lower_bound" = lower_bound, "upper_bound" = upper_bound)

means <- c()
lower_bound <- c()
upper_bound <- c()

for (j in 1:N) {
  bootData <- replicate (n=B, expr = rpois (n, lambda_hat))
  means<-colMeans(bootData)
  lower_bound[j] = quantile (means, 0.025)
  upper_bound[j] = quantile (means, 0.975)
}
ci_3 <- data.frame ("lower_bound" = lower_bound, "upper_bound" = upper_bound)

 means <- c()
 lower_bound <- c()
 upper_bound <- c()
 
 for (j in 1:N) { 
   lambda_vec <- rep(NA, B)
   for(i in 1:B){
     sel <- sample(1:n, n, replace=TRUE)
     bootstrap_x <- data[sel]
     lambda_vec[i] <- mean(bootstrap_x)  }
   lower_bound[j] <- quantile(lambda_vec, 0.025)
   upper_bound[j] <- quantile(lambda_vec, 0.975)}
 ci_4 <- data.frame ("lower_bound" = lower_bound, "upper_bound" = upper_bound)
```

## 3. Summary about different method execution

Calculate proportion of times each method contains the true value of the mean = 1.
```{r}
ci_1_prop <- nrow (subset (ci_1, ci_1$lower_bound<=1 & ci_1$upper_bound >=1))/nrow(ci_1)
ci_2_prop <- nrow (subset (ci_2, ci_2$lower_bound<=1 & ci_2$upper_bound >=1))/nrow(ci_2)
ci_3_prop <- nrow (subset (ci_3, ci_3$lower_bound<=1 & ci_3$upper_bound >=1))/nrow(ci_3)
ci_4_prop <- nrow (subset (ci_4, ci_4$lower_bound<=1 & ci_4$upper_bound >=1))/nrow(ci_4)
```

Calculate the average length of the CIs created by the procedure.
```{r}
ci_1_avg <- mean (ci_1$upper_bound - ci_1$lower_boun)
ci_2_avg <- mean (ci_2$upper_bound - ci_2$lower_boun)
ci_3_avg <- mean (ci_3$upper_bound - ci_3$lower_boun)
ci_4_avg <- mean (ci_4$upper_bound - ci_4$lower_boun)

dt <- data.frame (c ("Approximate 1", "Approximate 2", "Parametric Bootstrap", "Nonparametric Bootstrap"), 
                  c(ci_1_prop, ci_2_prop, ci_3_prop, ci_4_prop), 
                  c (ci_1_avg, ci_2_avg, ci_3_avg, ci_4_avg))
colnames(dt) <- c ("Method", "Prop Containing", "Avg Length")
kable(dt)
```

## 4. Repeat steps above for diffrent sample sizes 

## 5. Advantages and disadvantages of different types of interval 

## 6. Apply the 4 different methods to the horse-kick death data and report the 95% confidence intervals obtained by each method.

